
---
title: "Practice Questions for Exam 2"
output: html_document
---

```{r, include=FALSE}
library(dplyr)     
library(ggplot2)    
    
```




### Question 1
Load the `mtcars` dataset. Use `dplyr` to filter rows where the `mpg` is greater than 25 and display the first 5 rows.

```{r}
data(mtcars) 
data_mtcars <- mtcars %>% filter(mtcars$mpg > 25)
head(data_mtcars, 5)
```

### Question 2
Using the `iris` dataset, group the data by `Species` and calculate the mean `Sepal.Length` for each group.

```{r}
data(iris)
iris_Species <- iris %>% group_by(Species) %>% summarise(Sepal.Length_mean = mean(Sepal.Length))
iris_Species
```

### Question 3
Select only the `cyl` and `hp` columns from the `mtcars` dataset and arrange the rows in descending order of `hp`.

```{r}
data(mtcars)
mtcars_cyl_and_hp <- mtcars %>% select(cyl, hp) %>% arrange(desc(hp))
mtcars_cyl_and_hp
```

### Question 4
Use `mutate` on the `mtcars` dataset to create a new column `weight_kg` by converting the `wt` column (in 1000 lbs) into kilograms.

```{r}
data(mtcars)
mtcars_lbs_to_kg <- mtcars %>% mutate(weight_kg = wt * 0.45)
mtcars_lbs_to_kg
```

### Question 5
Filter the `diamonds` dataset to show only diamonds with a `carat` greater than 2.5 and a `price` less than 10,000.

```{r}
data(diamonds)
diamonds_filtered <- diamonds %>% filter(carat > 2.5, price < 10000)
diamonds_filtered
```

### Question 6
Write a pipeline to:
- Filter rows in `iris` where `Sepal.Width` > 3.
- Select the columns `Sepal.Length` and `Species`.
- Arrange the resulting rows by `Sepal.Length` in ascending order.

```{r}
data(iris)
iris_filtered <- iris %>% filter(Sepal.Width > 3) %>% select(Sepal.Length, Species) %>% arrange(Sepal.Length)
iris_filtered
```

### Question 7
Summarize the `mtcars` dataset to find the average horsepower (`hp`) for each unique number of cylinders (`cyl`).

```{r}
data(mtcars)
mtcars_avg <- mtcars %>% summarise(Mean_hp = mean(hp), Mean)
```

### Question 8
Add a new column to the `iris` dataset that categorizes `Petal.Length` as "Short" (< 2), "Medium" (2-4), or "Long" (> 4). Hint: use the case_when function which lets you specific that when a value in a column meets a certain condition, it takes on a certain value. Ex. case_when(Petal.Length == 2 ~ "Short")

```{r}
data(iris)
```


### Question 9
Compute the Euclidean distances between rows of the `mtcars` dataset using only the columns `mpg`, `disp`, and `hp`.

```{r}
data(mtcars)
```

### Question 10
Using the `iris` dataset, calculate the Manhattan distances between the first 10 rows using `Sepal.Length` and `Petal.Width`.

```{r}
data(iris)
```

### Question 11
Using the `mtcars` dataset's `mpg` and `hp` columns compute the Euclidean distances.

```{r}
data(mtcars)
```


### Question 12
Perform hierarchical clustering on the `iris` dataset using only the `Sepal.Length` and `Sepal.Width` columns. Use the `complete` linkage method and plot the resulting dendrogram.

```{r}
data(iris)
iris_clusters <-hclust(dist(as.matrix(iris[, c("Sepal.Length", "Sepal.Width")])), method = "complete")
iris_clusters
plot(iris_clusters)
```

### Question 13
Using the `mtcars` dataset, cluster the cars based on the `mpg`, `hp`, and `wt` columns. Use the average linkage method.

```{r}
data(mtcars)
mtcars_clusters <- hclust(dist(as.matrix(mtcars[, c("mpg", "hp", "wt")])), method = "average")
mtcars_clusters
plot(mtcars_clusters)
```

### Question 14
Write R code to put rectangles around 3 clusters in the dendrogram from clustering the `iris` dataset.

```{r}
data(iris)
iris_clusters <-hclust(dist(as.matrix(iris[, c("Sepal.Length", "Sepal.Width")])), method = "complete")
plot(iris_clusters)
rect.hclust(iris_clusters, k = 3, border = "red")

```

### Question 15
Write R code to extract the cluster assignments for 3 clusters from a dendrogram created using the `iris` dataset. Hint: Use cuttree.

```{r}
data(iris)
```

### Question 16
Perform k-means clustering on the `iris` dataset using `Petal.Length` and `Petal.Width`. Use `k = 3` and summarize the number of data points in each cluster.

```{r}
data(iris)
```

### Question 17
Using the `mtcars` dataset, determine the optimal number of clusters (`k`) using the Elbow Method for columns `mpg`, `hp`, and `wt`.

```{r}
data(mtcars)
wss <- numeric(8)
for (k in 1:8) {
  wss[k] <- kmeans(as.matrix(mtcars[, c("mpg", "hp", "wt")]), centers = k, nstart = 25)$tot.withinss
}
plot(1:8, wss, type = "b", xlab = "Number of Clusters", ylab = "Within groups sum of squares")

```

### Question 18
Plot the k-means clusters from the `iris` dataset (using `Petal.Length` and `Petal.Width`) on a scatterplot, coloring the points by their cluster assignments.

```{r}
data(iris)

```

### Question 19
Explain how the initial random centroids in k-means can impact the final cluster assignments. Perform k-means clustering on the `mtcars` dataset twice. Only use the mpg, hp, and wt columns.

```{r}
data(mtcars)

set.seed(123)
```

### Question 20
Compare the performance of k-means clustering and hierarchical clustering on the `mtcars` dataset based on columns `mpg`, `hp`, and `wt`. Do they broadly align in the placing the same items in the same clusters?

```{r}
data(mtcars)

```
